Group polarization replications may be marred by high false discovery rates

### 2024-12-30 GroupPolarizationStatmod/paper/ms.tex
eoy goalz

Our analysis estimates how the probability that a published detection of group
polarization is actually a false detection. We first identified latent opinion
parameters that yield spurious group polarization in an asymptotic $N\to\infty$
model. This gives us our first suggestion that experimental conditions may be
plausibly false. Briefly, we identified latent opinion distribution parameters
that generated spurious group polarization in 54 out of 57 experimental
conditions. In other words, 95\% of all group polarization detections are
plausibly spurious due to ceiling artifacts from the continuous latent
opinions-ordinal measurement experimental design. We then analyze family-wise
error and false detection rates for each experimental condition and aggregated 
up to the study level and across all studies and experimental conditions. 
Family-wise error rates varied from a minimum of 0.097 for the
\texttt{Feminists-Experimental} condition from Myers (1975)~\cite{Myers1975a}
and a maximum of 0.67 for the \texttt{COSprings-CivilUnions} condition from
Schkade, Sunstein, and Hastie (2010)~\cite{Schkade2010}. The median family-wise error
rate is 0.28 for Krizan and Baron's experimental condition
\texttt{NoOutgroupScenario1}. We close by inverting our procedure of calculating
family-wise error rates to calculating what $d^*$ would be required in each of the 54
experimental conditions to reduce the family-wise error rate to $\alpha = 0.05$.

@@

\clearpage

\begin{table}[ht]
\rotatebox{90} {
  \input{tex/quantile_table.tex}
}
\end{table}

@@

The $N\to\infty$ model enabled us to identify pre- and post-deliberation latent
opinion distribution parameters that caused spurious group polarization, denoted
$\beta_e$ for experimental condition $e$. Using these $\beta$ we drew 
$N_\mathrm{emp}$ samples from the pre- and post-deliberation latent opinion 
distributions, then binned them (Section~\ref{subsec:false-detections}) 
to simulate participants reporting their pre- and post-deliberation opinions. 
To estimate optimistic family-wise error and false detection rates, we fit an
ordered probit model to the simulated observations and obtained estimates of
the group polarization effect size, $d$ (Equation~\ref{eq:cohens}).
If the effect size was greater than a significance value, $d > d^*$, then there
was a detection $D$ of group polarization, even though by design the simulated latent mean
group opinion was constant, i.e., there in fact was no effect, written $\neg E$.
The family-wise error rate is the calculation of $\Pr(D|\neg E)$ as the frequency with which
$d > d^*$ out of 1000 simulation trials. The false detection rate is the family-wise
error rate scaled using the base rate of effects and the statistical power
(Equation~\ref{eq:fdr}). We test three difference significance values,
$d = 0.2,~0.5$ and $0.8$, corresponding to what Cohen prescribed to use for
``low'', ``medium'', and ``high'' significance, respectively.

We calculated the false detection rate across experimental conditions where
we identified parameters that generate spurious group polarization in the 
large-$N$ model (Figure~\ref{fig:fdr_experiments}). We then aggregated over
experimental conditions within each study to understand the false detection rate
across the ten studies (Figure~\ref{fig:fdr_studies}). Across the identified
spurious conditions, we calculated a mean false detection rate of 0.74 with
a significance level of $d^* = 0.5$. We also
examined the mean false detection rates across studies if we set a highly
optimistic estimate of the family-wise error rate, setting $\alpha = 0.05$. 
This only reduces the false discovery rate to 0.71, but does slightly change the
rankings of which study had the highest mean false discovery rates, with 
Schkade, et al., (2010) going from highest mean false discovery rate to 
second-lowest false discovery rate. The minimum false discovery rate in a study
was Myers (1975), with a false discovery rate of 0.58, meaning that probability
a group polarization detection in the best-designed study is greater than one in
two.

>>>>

Our analysis estimates how the probability that a published detection of group
polarization is actually a false detection. We first identified latent opinion
parameters that yield spurious group polarization in an asymptotic $N\to\infty$
model. This gives us our first suggestion that experimental conditions may be
plausibly false. Briefly, we identified latent opinion distribution parameters
that generated spurious group polarization in 54 out of 57 experimental
conditions. In other words, 95\% of all group polarization detections are
plausibly spurious due to ceiling artifacts from the continuous latent
opinions-ordinal measurement experimental design. We then analyze the family-wise
error ($\alpha$) and false detection rate (FDR) 
for each experimental condition and aggregated 
up to the study level and across all studies and experimental conditions. 
The lower twentieth percentile of $\alpha$
We close by showing that $d^*$ must be greater than the ``high'' effect size
$d^*=0.8$ to limit family-wise error rates to $\alpha \leq 0.05$ in all but six
experimental conditions.

@@

\begin{table}[ht]
  \caption{yo}
  \label{tab:quantiles}
\rotatebox{90} {\textbf{Quantiles for the two error-rate measures under different
  significance values.}\mt{Not sure about the best way to format; probably don't
  need this in the main text, but could be valuable in supplement?}}
  \input{tex/quantile_table.tex}

\end{table}

@@

We simulated trials from each experimental condition, $e$, 
by sampling and binning pre- and
post-deliberation latent opinions from distributions with parameters identified
with the $N\to\infty$ model as causing plausibly spurious group
polarization. For each $e$,
we drew a number of samples equal to the number of participants in 
the original experiments, $N_\mathrm{emp}$. To estimate
family-wise error and false detection rates, we fit an ordered probit
model to the simulated observations and obtained estimates of the group polarization
effect size, $d$ (Equation~\ref{eq:cohens}). If the effect size was greater than a
significance value, $d > d^*$, then by definition there was a detection $D$ of group 
polarization,
even though there was no effect by design, written $\neg E$. 
The family-wise error rate is $\Pr(D|\neg E)$, calculated as the frequency with which $d > d^*$ out of 1000
simulation trials. The false detection rate is the family-wise error rate scaled
using the base rate of effects and the statistical power (Equation~\ref{eq:fdr}). We
test three difference significance values, $d = 0.2,~0.5$ and $0.8$, corresponding
to what Cohen prescribed to use for ``low'', ``medium'', and ``high'' significance.

Family-wise error rates (Figure~\ref{fig:fwer_fdr_synthesis} and
Table~\ref{tab:quantiles}) varied from a minimum of one in ten for the
\texttt{Feminists-Experimental} condition from Myers (1975)~\cite{Myers1975a}
to a maximum of two in three for the \texttt{COSprings-CivilUnions} condition from
Schkade, Sunstein, and Hastie (2010)~\cite{Schkade2010}. The median family-wise error
rate is 0.28 for Krizan and Baron's (2007) experimental condition
\texttt{NoOutgroupScenario1}~\cite{Krizan2007}.
